---
title: "AL_vars"
author: "Elizabeth Baach"
output:
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Loading packages
```{r message=FALSE, warning=FALSE, paged.print=FALSE}

library(dplyr)
library(tidyverse)
library(rFIA)
```
### Loading in FIA tables
(these are available online at the FIA datamart)
```{r}

AL <- readFIA('C:/Users/elbaa/OneDrive/Desktop/gitMSDATA/AL_data', inMemory = FALSE)

AL_TREE <- readFIA(dir = 'C:/Users/elbaa/OneDrive/Desktop/gitMSDATA/AL_data', tables = 'TREE', inMemory = TRUE)
AL_COND <- readFIA(dir = 'C:/Users/elbaa/OneDrive/Desktop/gitMSDATA/AL_data', tables = 'COND', inMemory = TRUE)
AL_PLOT <- readFIA(dir = 'C:/Users/elbaa/OneDrive/Desktop/gitMSDATA/AL_data', tables = 'PLOT', inMemory = TRUE)
```

### Filtering the tables
```{r}

# filtering TREE table
tree_tab <- AL_TREE$TREE
tree_tab <- select(tree_tab,c(CN, PLT_CN, PREV_TRE_CN, INVYR, COUNTYCD, PLOT, SUBP, CONDID, STATUSCD, SPCD, DIA, HT, SPGRPCD,CR, CCLCD,SPGRPCD,UNITCD))
tree_tab <- tree_tab %>% mutate(diam = (2.54 * tree_tab$DIA ))
tree_tab<- tree_tab %>% select(-c(DIA))
tree_tab$DIA<- tree_tab$diam
tree_tab<- tree_tab %>% select(-c(diam))
 
# filtering PLOT table
plot_tab <- AL_PLOT$PLOT
plot_tab<- select(plot_tab, c(PLOT, INVYR, COUNTYCD, LAT, LON, ELEV, RDDISTCD, ECOSUBCD))
plot_tab$COUNTY_PLOT <- paste(plot_tab$COUNTYCD,plot_tab$PLOT, sep="_")
plot_tab<- plot_tab %>% filter(INVYR>=2009)
 
# filtering COND table
cond_tab <- AL_COND$COND
cond_tab<- select(cond_tab, c(CN, PLT_CN, INVYR, COUNTYCD, PLOT, FORTYPCD, STDAGE, STDORGCD, DSTRBCD1,TRTCD1, SITECLCD, SICOND,PHYSCLCD, FIRE_SRS, GRAZING_SRS, OWNCD, STAND_STRUCTURE_SRS, BALIVE,SISP,SLOPE,ASPECT))
cond_tab<- cond_tab %>% filter(INVYR>=2009)
cond_tab<- cond_tab %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
cond_tab$COUNTY_PLOT <- paste(cond_tab$COUNTYCD,cond_tab$PLOT, sep="_")

```

### Subsetting by the year
For this study we used data from 2009 to 2019 only
```{r}
#tree
tree09 <- tree_tab %>% subset(INVYR==2009)
tree10 <- tree_tab %>% subset(INVYR==2010)
tree11 <- tree_tab %>% subset(INVYR==2011)
tree12 <- tree_tab %>% subset(INVYR==2012)
tree13 <- tree_tab %>% subset(INVYR==2013)
tree14 <- tree_tab %>% subset(INVYR==2014)
tree15 <- tree_tab %>% subset(INVYR==2015)
tree16 <- tree_tab %>% subset(INVYR==2016)
tree17 <- tree_tab %>% subset(INVYR==2017)
tree18 <- tree_tab %>% subset(INVYR==2018)
tree19 <- tree_tab %>% subset(INVYR==2019)
#plot
plot09<- plot_tab %>% subset(INVYR==2009)
plot10<- plot_tab %>% subset(INVYR==2010)
plot11<- plot_tab %>% subset(INVYR==2011)
plot12<- plot_tab %>% subset(INVYR==2012)
plot13<- plot_tab %>% subset(INVYR==2013)
plot14<- plot_tab %>% subset(INVYR==2014)
plot15<- plot_tab %>% subset(INVYR==2015)
plot16<- plot_tab %>% subset(INVYR==2016)
plot17<- plot_tab %>% subset(INVYR==2017)
plot18<- plot_tab %>% subset(INVYR==2018)
plot19<- plot_tab %>% subset(INVYR==2019)
#cond
cond09<- cond_tab %>% subset(INVYR==2009)
cond10<- cond_tab %>% subset(INVYR==2010)
cond11<- cond_tab %>% subset(INVYR==2011)
cond12<- cond_tab %>% subset(INVYR==2012)
cond13<- cond_tab %>% subset(INVYR==2013)
cond14<- cond_tab %>% subset(INVYR==2014)
cond15<- cond_tab %>% subset(INVYR==2015)
cond16<- cond_tab %>% subset(INVYR==2016)
cond17<- cond_tab %>% subset(INVYR==2017)
cond18<- cond_tab %>% subset(INVYR==2018)
cond19<- cond_tab %>% subset(INVYR==2019)
```

### finding remearsured plots
this was orginally done for all year combinations-- remaining code show those with remeasurement data availablie
```{r}
checkdat1409 <- subset(tree14, tree14$PREV_TRE_CN %in% tree09$CN)
checkdat1409 <- subset(checkdat1409, select = -c(CN))
checkdat1409$CN <-     checkdat1409$PREV_TRE_CN
checkdat1409 <- subset(checkdat1409, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1509 <- subset(tree15, tree15$PREV_TRE_CN %in% tree09$CN)
checkdat1509 <- subset(checkdat1509, select = -c(CN))
checkdat1509$CN <-     checkdat1509$PREV_TRE_CN
checkdat1509 <- subset(checkdat1509, select = -c(PREV_TRE_CN, PLT_CN))


checkdat1609 <- subset(tree16, tree16$PREV_TRE_CN %in% tree09$CN)
checkdat1609 <- subset(checkdat1609, select = -c(CN))
checkdat1609$CN <- checkdat1609$PREV_TRE_CN
checkdat1609 <- subset(checkdat1609, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1709 <- subset(tree17, tree17$PREV_TRE_CN %in% tree09$CN)
checkdat1709 <- subset(checkdat1709, select = -c(CN))
checkdat1709$CN <-     checkdat1709$PREV_TRE_CN
checkdat1709 <- subset(checkdat1709, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1809 <- subset(tree18, tree18$PREV_TRE_CN %in% tree09$CN)
checkdat1809 <- subset(checkdat1809, select = -c(CN))
checkdat1809$CN <-     checkdat1809$PREV_TRE_CN
checkdat1809 <- subset(checkdat1809, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1510 <- subset(tree15, tree15$PREV_TRE_CN %in% tree10$CN)
checkdat1510 <- subset(checkdat1510, select = -c(CN))
checkdat1510$CN <-     checkdat1510$PREV_TRE_CN
checkdat1510 <- subset(checkdat1510, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1610 <- subset(tree16, tree16$PREV_TRE_CN %in% tree10$CN)
checkdat1610 <- subset(checkdat1610, select = -c(CN))
checkdat1610$CN <-     checkdat1610$PREV_TRE_CN
checkdat1610 <- subset(checkdat1610, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1710 <- subset(tree17, tree17$PREV_TRE_CN %in% tree10$CN)
checkdat1710 <- subset(checkdat1710, select = -c(CN))
checkdat1710$CN <-     checkdat1710$PREV_TRE_CN
checkdat1710 <- subset(checkdat1710, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1810 <- subset(tree18, tree18$PREV_TRE_CN %in% tree10$CN)
checkdat1810 <- subset(checkdat1810, select = -c(CN))
checkdat1810$CN <-     checkdat1810$PREV_TRE_CN
checkdat1810 <- subset(checkdat1810, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1511 <- subset(tree15, tree15$PREV_TRE_CN %in% tree11$CN)
checkdat1511 <- subset(checkdat1511, select = -c(CN))
checkdat1511$CN <-     checkdat1511$PREV_TRE_CN
checkdat1511 <- subset(checkdat1511, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1611 <- subset(tree16, tree16$PREV_TRE_CN %in% tree11$CN)
checkdat1611 <- subset(checkdat1611, select = -c(CN))
checkdat1611$CN <-     checkdat1611$PREV_TRE_CN
checkdat1611 <- subset(checkdat1611, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1711 <- subset(tree17, tree17$PREV_TRE_CN %in% tree11$CN)
checkdat1711 <- subset(checkdat1711, select = -c(CN))
checkdat1711$CN <-     checkdat1711$PREV_TRE_CN
checkdat1711 <- subset(checkdat1711, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1811 <- subset(tree18, tree18$PREV_TRE_CN %in% tree11$CN)
checkdat1811 <- subset(checkdat1811, select = -c(CN))
checkdat1811$CN <-     checkdat1811$PREV_TRE_CN
checkdat1811 <- subset(checkdat1811, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1911 <- subset(tree19, tree19$PREV_TRE_CN %in% tree11$CN)
checkdat1911 <- subset(checkdat1911, select = -c(CN))
checkdat1911$CN <-     checkdat1911$PREV_TRE_CN
checkdat1911 <- subset(checkdat1911, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1612 <- subset(tree16, tree16$PREV_TRE_CN %in% tree12$CN)
checkdat1612 <- subset(checkdat1612, select = -c(CN))
checkdat1612$CN <-     checkdat1612$PREV_TRE_CN
checkdat1612 <- subset(checkdat1612, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1712 <- subset(tree17, tree17$PREV_TRE_CN %in% tree12$CN)
checkdat1712 <- subset(checkdat1712, select = -c(CN))
checkdat1712$CN <-     checkdat1712$PREV_TRE_CN
checkdat1712 <- subset(checkdat1712, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1812 <- subset(tree18, tree18$PREV_TRE_CN %in% tree12$CN)
checkdat1812 <- subset(checkdat1812, select = -c(CN))
checkdat1812$CN <-     checkdat1812$PREV_TRE_CN
checkdat1812 <- subset(checkdat1812, select = -c(PREV_TRE_CN, PLT_CN))

checkdat1912 <- subset(tree19, tree19$PREV_TRE_CN %in% tree12$CN)
checkdat1912 <- subset(checkdat1912, select = -c(CN))
checkdat1912$CN <-     checkdat1912$PREV_TRE_CN
checkdat1912 <- subset(checkdat1912, select = -c(PREV_TRE_CN, PLT_CN))


#and the reverse
checkdat0914 <- subset(tree09, tree09$CN %in%  tree14$PREV_TRE_CN)
checkdat0914 <- subset(checkdat0914, select = -c(PREV_TRE_CN))

checkdat0915 <- subset(tree09, tree09$CN %in%  tree15$PREV_TRE_CN)
checkdat0915 <- subset(checkdat0915, select = -c(PREV_TRE_CN))

checkdat0916 <- subset(tree09, tree09$CN %in%  tree16$PREV_TRE_CN)
checkdat0916 <- subset(checkdat0916, select = -c(PREV_TRE_CN))

checkdat0917 <- subset(tree09, tree09$CN %in%  tree17$PREV_TRE_CN)
checkdat0917 <- subset(checkdat0917, select = -c(PREV_TRE_CN))

checkdat0918 <- subset(tree09, tree09$CN %in%  tree18$PREV_TRE_CN)
checkdat0918 <- subset(checkdat0918, select = -c(PREV_TRE_CN))

checkdat1015 <- subset(tree10, tree10$CN %in%  tree15$PREV_TRE_CN)
checkdat1015 <- subset(checkdat1015, select = -c(PREV_TRE_CN))

checkdat1016 <- subset(tree10, tree10$CN %in%  tree16$PREV_TRE_CN)
checkdat1016 <- subset(checkdat1016, select = -c(PREV_TRE_CN))

checkdat1017 <- subset(tree10, tree10$CN %in%  tree17$PREV_TRE_CN)
checkdat1017 <- subset(checkdat1017, select = -c(PREV_TRE_CN))

checkdat1018 <- subset(tree10, tree10$CN %in%  tree18$PREV_TRE_CN)
checkdat1018 <- subset(checkdat1018, select = -c(PREV_TRE_CN))

checkdat1115 <- subset(tree11, tree11$CN %in%  tree15$PREV_TRE_CN)
checkdat1115 <- subset(checkdat1115, select = -c(PREV_TRE_CN))

checkdat1116 <- subset(tree11, tree11$CN %in%  tree16$PREV_TRE_CN)
checkdat1116 <- subset(checkdat1116, select = -c(PREV_TRE_CN))

checkdat1117 <- subset(tree11, tree11$CN %in%  tree17$PREV_TRE_CN)
checkdat1117 <- subset(checkdat1117, select = -c(PREV_TRE_CN))

checkdat1118 <- subset(tree11, tree11$CN %in%  tree18$PREV_TRE_CN)
checkdat1118 <- subset(checkdat1118, select = -c(PREV_TRE_CN))

checkdat1119 <- subset(tree11, tree11$CN %in%  tree19$PREV_TRE_CN)
checkdat1119 <- subset(checkdat1119, select = -c(PREV_TRE_CN))

checkdat1216 <- subset(tree12, tree12$CN %in%  tree16$PREV_TRE_CN)
checkdat1216 <- subset(checkdat1216, select = -c(PREV_TRE_CN))

checkdat1217 <- subset(tree12, tree12$CN %in%  tree17$PREV_TRE_CN)
checkdat1217 <- subset(checkdat1217, select = -c(PREV_TRE_CN))

checkdat1218 <- subset(tree12, tree12$CN %in%  tree18$PREV_TRE_CN)
checkdat1218 <- subset(checkdat1218, select = -c(PREV_TRE_CN))

checkdat1219 <- subset(tree12, tree12$CN %in%  tree19$PREV_TRE_CN)
checkdat1219 <- subset(checkdat1219, select = -c(PREV_TRE_CN))
```

### Merging remeasured plot data
```{r}
#2009 to 2014
data0914<- checkdat1409
data0914$DIA09 <- checkdat0914$DIA[match(checkdat1409$CN,checkdat0914$CN)]
data0914$DIA14<- data0914$DIA
data0914<- data0914 %>% select(-c(DIA ))
data0914$HT09<- checkdat0914$HT[match(checkdat1409$CN, checkdat0914$CN)]
data0914$HT14<- data0914$HT
data0914<- data0914 %>% select(-c(HT, INVYR))
data0914<- data0914 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data0914$COUNTY_PLOT <- paste(data0914$COUNTYCD,data0914$PLOT, sep="_")
data0914$BA<- pi*(.5*data0914$DIA09)^2
data0914$STDAGE <- cond09$STDAGE[match(data0914$COUNTY_PLOT, cond09$COUNTY_PLOT)]

#2009 to 2015
data0915<- checkdat1509
data0915$DIA09 <- checkdat0915$DIA[match(checkdat1509$CN,checkdat0915$CN)]
data0915$DIA15<- data0915$DIA
data0915<- data0915 %>% select(-c(DIA ))
data0915$HT09<- checkdat0915$HT[match(checkdat1509$CN, checkdat0915$CN)]
data0915$HT15<- data0915$HT
data0915<- data0915 %>% select(-c(HT, INVYR))
data0915<- data0915 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data0915$COUNTY_PLOT <- paste(data0915$COUNTYCD,data0915$PLOT, sep="_")
data0915$BA<- pi*(.5*data0915$DIA09)^2
data0915$STDAGE <- cond09$STDAGE[match(data0915$COUNTY_PLOT, cond09$COUNTY_PLOT)]

#2009 to 2016
data0916<- checkdat1609
data0916$DIA09 <- checkdat0916$DIA[match(checkdat1609$CN,checkdat0916$CN)]
data0916$DIA16<- data0916$DIA
data0916<- data0916 %>% select(-c(DIA))
data0916$HT09<- checkdat0916$HT[match(checkdat1609$CN, checkdat0916$CN)]
data0916$HT16<- data0916$HT
data0916<- data0916 %>% select(-c(HT, INVYR))
data0916<- data0916 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
#now i will make a new col that has county and plot
data0916$COUNTY_PLOT <- paste(data0916$COUNTYCD,data0916$PLOT, sep="_")
data0916$CNTY_PLT_SPCD<-paste(data0916$COUNTY_PLOT,data0916$SPCD,sep="#")
#adding BA and stand age using the first of the years
data0916$BA<- pi*(.5*data0916$DIA09)^2
data0916$STDAGE <- cond09$STDAGE[match(data0916$COUNTY_PLOT, cond09$COUNTY_PLOT)]

#2009 to 2017
data0917<- checkdat1709
data0917$DIA09 <- checkdat0917$DIA[match(checkdat1709$CN,checkdat0917$CN)]
data0917$DIA17<- data0917$DIA
data0917<- data0917 %>% select(-c(DIA ))
data0917$HT09<- checkdat0917$HT[match(checkdat1709$CN, checkdat0917$CN)]
data0917$HT17<- data0917$HT
data0917<- data0917 %>% select(-c(HT, INVYR))
data0917<- data0917 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data0917$COUNTY_PLOT <- paste(data0917$COUNTYCD,data0917$PLOT, sep="_")
data0917$BA<- pi*(.5*data0917$DIA09)^2
data0917$STDAGE <- cond09$STDAGE[match(data0917$COUNTY_PLOT, cond09$COUNTY_PLOT)]

#2009 to 2018
data0918<- checkdat1809
data0918$DIA09 <- checkdat0918$DIA[match(checkdat1809$CN,checkdat0918$CN)]
data0918$DIA18<- data0918$DIA
data0918<- data0918 %>% select(-c(DIA ))
data0918$HT09<- checkdat0918$HT[match(checkdat1809$CN, checkdat0918$CN)]
data0918$HT18<- data0918$HT
data0918<- data0918 %>% select(-c(HT, INVYR))
data0918<- data0918 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data0918$COUNTY_PLOT <- paste(data0918$COUNTYCD,data0918$PLOT, sep="_")
data0918$BA<- pi*(.5*data0918$DIA09)^2
data0918$STDAGE <- cond09$STDAGE[match(data0918$COUNTY_PLOT, cond09$COUNTY_PLOT)]

#2010 to 2015
data1015<-checkdat1510
data1015$DIA10 <- checkdat1015$DIA[match(checkdat1510$CN,checkdat1015$CN)]
data1015$DIA15<- data1015$DIA
data1015<- data1015 %>% select(-c(DIA ))
data1015$HT10<- checkdat1015$HT[match(checkdat1510$CN, checkdat1015$CN)]
data1015$HT16<- data1015$HT
data1015<- data1015 %>% select(-c(HT, INVYR))
data1015<- data1015 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1015$COUNTY_PLOT <- paste(data1015$COUNTYCD,data1015$PLOT, sep="_")
data1015$BA<- pi*(.5*data1015$DIA10)^2
data1015$STDAGE <- cond10$STDAGE[match(data1015$COUNTY_PLOT, cond10$COUNTY_PLOT)]

#2010 to 2016
data1016<-checkdat1610
data1016$DIA10 <- checkdat1016$DIA[match(checkdat1610$CN,checkdat1016$CN)]
data1016$DIA16<- data1016$DIA
data1016<- data1016 %>% select(-c(DIA ))
data1016$HT10<- checkdat1016$HT[match(checkdat1610$CN, checkdat1016$CN)]
data1016$HT16<- data1016$HT
data1016<- data1016 %>% select(-c(HT, INVYR))
data1016<- data1016 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1016$COUNTY_PLOT <- paste(data1016$COUNTYCD,data1016$PLOT, sep="_")
data1016$BA<- pi*(.5*data1016$DIA10)^2
data1016$STDAGE <- cond10$STDAGE[match(data1016$COUNTY_PLOT, cond10$COUNTY_PLOT)]

#2010 to 2017
data1017<-checkdat1710
data1017$DIA10 <- checkdat1017$DIA[match(checkdat1710$CN,checkdat1017$CN)]
data1017$DIA17<- data1017$DIA
data1017<- data1017 %>% select(-c(DIA ))
data1017$HT10<- checkdat1017$HT[match(checkdat1710$CN, checkdat1017$CN)]
data1017$HT17<- data1017$HT
data1017<- data1017 %>% select(-c(HT, INVYR))
data1017<- data1017 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1017$COUNTY_PLOT <- paste(data1017$COUNTYCD,data1017$PLOT, sep="_")
data1017$BA<- pi*(.5*data1017$DIA10)^2
data1017$STDAGE <- cond10$STDAGE[match(data1017$COUNTY_PLOT, cond10$COUNTY_PLOT)]

#2010 to 2018
data1018<-checkdat1810
data1018$DIA10 <- checkdat1018$DIA[match(checkdat1810$CN,checkdat1018$CN)]
data1018$DIA18<- data1018$DIA
data1018<- data1018 %>% select(-c(DIA ))
data1018$HT10<- checkdat1018$HT[match(checkdat1810$CN, checkdat1018$CN)]
data1018$HT18<- data1018$HT
data1018<- data1018 %>% select(-c(HT, INVYR))
data1018<- data1018 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1018$COUNTY_PLOT <- paste(data1018$COUNTYCD,data1018$PLOT, sep="_")
data1018$BA<- pi*(.5*data1018$DIA10)^2
data1018$STDAGE <- cond10$STDAGE[match(data1018$COUNTY_PLOT, cond10$COUNTY_PLOT)]

#2011 to 2015
data1115<-checkdat1511
data1115$DIA11 <- checkdat1115$DIA[match(checkdat1511$CN,checkdat1115$CN)]
data1115$DIA15<- data1115$DIA
data1115<- data1115 %>% select(-c(DIA ))
data1115$HT11<- checkdat1115$HT[match(checkdat1511$CN, checkdat1115$CN)]
data1115$HT15<- data1115$HT
data1115<- data1115 %>% select(-c(HT, INVYR))
data1115<- data1115 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1115$COUNTY_PLOT <- paste(data1115$COUNTYCD,data1115$PLOT, sep="_")
data1115$BA<- pi*(.5*data1115$DIA11)^2
data1115$STDAGE <- cond11$STDAGE[match(data1115$COUNTY_PLOT, cond11$COUNTY_PLOT)]

#2011 to 2016
data1116<-checkdat1611
data1116$DIA11 <- checkdat1116$DIA[match(checkdat1611$CN,checkdat1116$CN)]
data1116$DIA16<- data1116$DIA
data1116<- data1116 %>% select(-c(DIA ))
data1116$HT11<- checkdat1116$HT[match(checkdat1611$CN, checkdat1116$CN)]
data1116$HT16<- data1116$HT
data1116<- data1116 %>% select(-c(HT, INVYR))
data1116<- data1116 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1116$COUNTY_PLOT <- paste(data1116$COUNTYCD,data1116$PLOT, sep="_")
data1116$BA<- pi*(.5*data1116$DIA11)^2
data1116$STDAGE <- cond11$STDAGE[match(data1116$COUNTY_PLOT, cond11$COUNTY_PLOT)]

#2011 to 2017
data1117<-checkdat1711
data1117$DIA11 <- checkdat1117$DIA[match(checkdat1711$CN,checkdat1117$CN)]
data1117$DIA17<- data1117$DIA
data1117<- data1117 %>% select(-c(DIA ))
data1117$HT11<- checkdat1117$HT[match(checkdat1711$CN, checkdat1117$CN)]
data1117$HT17<- data1117$HT
data1117<- data1117 %>% select(-c(HT, INVYR))
data1117<- data1117 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1117$COUNTY_PLOT <- paste(data1117$COUNTYCD,data1117$PLOT, sep="_")
data1117$BA<- pi*(.5*data1117$DIA11)^2
data1117$STDAGE <- cond11$STDAGE[match(data1117$COUNTY_PLOT, cond11$COUNTY_PLOT)]

#2011 to 2018
data1118<-checkdat1811
data1118$DIA11 <- checkdat1118$DIA[match(checkdat1811$CN,checkdat1118$CN)]
data1118$DIA18<- data1118$DIA
data1118<- data1118 %>% select(-c(DIA ))
data1118$HT11<- checkdat1118$HT[match(checkdat1811$CN, checkdat1118$CN)]
data1118$HT18<- data1118$HT
data1118<- data1118 %>% select(-c(HT, INVYR))
data1118<- data1118 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1118$COUNTY_PLOT <- paste(data1118$COUNTYCD,data1118$PLOT, sep="_")
data1118$BA<- pi*(.5*data1118$DIA11)^2
data1118$STDAGE <- cond11$STDAGE[match(data1118$COUNTY_PLOT, cond11$COUNTY_PLOT)]

#2011 to 2019
data1119<-checkdat1911
data1119$DIA11 <- checkdat1119$DIA[match(checkdat1911$CN,checkdat1119$CN)]
data1119$DIA19<- data1119$DIA
data1119<- data1119 %>% select(-c(DIA ))
data1119$HT11<- checkdat1119$HT[match(checkdat1911$CN, checkdat1119$CN)]
data1119$HT19<- data1119$HT
data1119<- data1119 %>% select(-c(HT, INVYR))
data1119<- data1119 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1119$COUNTY_PLOT <- paste(data1119$COUNTYCD,data1119$PLOT, sep="_")
data1119$BA<- pi*(.5*data1119$DIA11)^2
data1119$STDAGE <- cond11$STDAGE[match(data1119$COUNTY_PLOT, cond11$COUNTY_PLOT)]

#2012 to 2016
data1216<-checkdat1612
data1216$DIA12 <- checkdat1216$DIA[match(checkdat1612$CN,checkdat1216$CN)]
data1216$DIA16<- data1216$DIA
data1216<- data1216 %>% select(-c(DIA ))
data1216$HT12<- checkdat1216$HT[match(checkdat1612$CN, checkdat1216$CN)]
data1216$HT16<- data1216$HT
data1216<- data1216 %>% select(-c(HT, INVYR))
data1216<- data1216 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1216$COUNTY_PLOT <- paste(data1216$COUNTYCD,data1216$PLOT, sep="_")
data1216$BA<- pi*(.5*data1216$DIA12)^2
data1216$STDAGE <- cond12$STDAGE[match(data1216$COUNTY_PLOT, cond12$COUNTY_PLOT)]

#2012 to 2017
data1217<-checkdat1712
data1217$DIA12 <- checkdat1217$DIA[match(checkdat1712$CN,checkdat1217$CN)]
data1217$DIA17<- data1217$DIA
data1217<- data1217 %>% select(-c(DIA ))
data1217$HT12<- checkdat1217$HT[match(checkdat1712$CN, checkdat1217$CN)]
data1217$HT17<- data1217$HT
data1217<- data1217 %>% select(-c(HT, INVYR))
data1217<- data1217 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1217$COUNTY_PLOT <- paste(data1217$COUNTYCD,data1217$PLOT, sep="_")
data1217$BA<- pi*(.5*data1217$DIA12)^2
data1217$STDAGE <- cond12$STDAGE[match(data1217$COUNTY_PLOT, cond12$COUNTY_PLOT)]

#2012 to 2018
data1218<-checkdat1812
data1218$DIA12 <- checkdat1218$DIA[match(checkdat1812$CN,checkdat1218$CN)]
data1218$DIA18<- data1218$DIA
data1218<- data1218 %>% select(-c(DIA ))
data1218$HT12<- checkdat1218$HT[match(checkdat1812$CN, checkdat1218$CN)]
data1218$HT18<- data1218$HT
data1218<- data1218 %>% select(-c(HT, INVYR))
data1218<- data1218 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1218$COUNTY_PLOT <- paste(data1218$COUNTYCD,data1218$PLOT, sep="_")
data1218$BA<- pi*(.5*data1218$DIA12)^2
data1218$STDAGE <- cond12$STDAGE[match(data1218$COUNTY_PLOT, cond12$COUNTY_PLOT)]

#2012 to 2019
data1219<-checkdat1912
data1219$DIA12 <- checkdat1219$DIA[match(checkdat1912$CN,checkdat1219$CN)]
data1219$DIA19<- data1219$DIA
data1219<- data1219 %>% select(-c(DIA ))
data1219$HT12<- checkdat1219$HT[match(checkdat1912$CN, checkdat1219$CN)]
data1219$HT19<- data1219$HT
data1219<- data1219 %>% select(-c(HT, INVYR))
data1219<- data1219 %>% group_by(COUNTYCD) %>% arrange(PLOT, .by_group=TRUE)
data1219$COUNTY_PLOT <- paste(data1219$COUNTYCD,data1219$PLOT, sep="_")
data1219$BA<- pi*(.5*data1219$DIA12)^2
data1219$STDAGE <- cond12$STDAGE[match(data1219$COUNTY_PLOT, cond12$COUNTY_PLOT)]
```

### Filtering for treatment and disturbance and natrual regen
```{r}
#2009-14
data0914$STDORGCD09 <- cond09$STDORGCD[match(data0914$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0914$STDORGCD14 <- cond14$STDORGCD[match(data0914$COUNTY_PLOT, cond14$COUNTY_PLOT)]
data0914<- data0914 %>% filter(STDORGCD09==0)
data0914<- data0914 %>% filter(STDORGCD14==0)
data0914<- data0914 %>% select(-c(STDORGCD09,STDORGCD14))

data0914$DSTRB09 <- cond09$DSTRBCD1[match(data0914$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0914$DSTRB14 <- cond14$DSTRBCD1[match(data0914$COUNTY_PLOT, cond14$COUNTY_PLOT)]
data0914<- data0914 %>% filter(DSTRB09==0)
data0914<- data0914 %>% filter(DSTRB14==0)
data0914<- data0914 %>% select(-c(DSTRB09, DSTRB14))

data0914$TRT09 <- cond09$TRTCD1[match(data0914$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0914$TRT14 <- cond14$TRTCD1[match(data0914$COUNTY_PLOT, cond14$COUNTY_PLOT)]
data0914<- data0914 %>% filter(TRT09==0)
data0914<- data0914 %>% filter(TRT14==0)
data0914<- data0914 %>% select(-c(TRT09, TRT14))

#2009-15
data0915$STDORGCD09 <- cond09$STDORGCD[match(data0915$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0915$STDORGCD15 <- cond15$STDORGCD[match(data0915$COUNTY_PLOT, cond15$COUNTY_PLOT)]
data0915<- data0915 %>% filter(STDORGCD09==0)
data0915<- data0915 %>% filter(STDORGCD15==0)
data0915<- data0915 %>% select(-c(STDORGCD09,STDORGCD15))

data0915$DSTRB09 <- cond09$DSTRBCD1[match(data0915$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0915$DSTRB15 <- cond15$DSTRBCD1[match(data0915$COUNTY_PLOT, cond15$COUNTY_PLOT)]
data0915<- data0915 %>% filter(DSTRB09==0)
data0915<- data0915 %>% filter(DSTRB15==0)
data0915<- data0915 %>% select(-c(DSTRB09, DSTRB15))

data0915$TRT09 <- cond09$TRTCD1[match(data0915$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0915$TRT15 <- cond15$TRTCD1[match(data0915$COUNTY_PLOT, cond15$COUNTY_PLOT)]
data0915<- data0915 %>% filter(TRT09==0)
data0915<- data0915 %>% filter(TRT15==0)
data0915<- data0915 %>% select(-c(TRT09, TRT15))

#2009-16
data0916$STDORGCD09 <- cond09$STDORGCD[match(data0916$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0916$STDORGCD16 <- cond16$STDORGCD[match(data0916$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data0916<- data0916 %>% filter(STDORGCD09==0)
data0916<- data0916 %>% filter(STDORGCD16==0)
data0916<- data0916 %>% select(-c(STDORGCD09,STDORGCD16))

data0916$DSTRB09 <- cond09$DSTRBCD1[match(data0916$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0916$DSTRB16 <- cond16$DSTRBCD1[match(data0916$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data0916<- data0916 %>% filter(DSTRB09==0)
data0916<- data0916 %>% filter(DSTRB16==0)
data0916<- data0916 %>% select(-c(DSTRB09, DSTRB16))

data0916$TRT09 <- cond09$TRTCD1[match(data0916$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0916$TRT16 <- cond16$TRTCD1[match(data0916$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data0916<- data0916 %>% filter(TRT09==0)
data0916<- data0916 %>% filter(TRT16==0)
data0916<- data0916 %>% select(-c(TRT09, TRT16))

#2009-17
data0917$STDORGCD09 <- cond09$STDORGCD[match(data0917$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0917$STDORGCD17 <- cond17$STDORGCD[match(data0917$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data0917<- data0917 %>% filter(STDORGCD09==0)
data0917<- data0917 %>% filter(STDORGCD17==0)
data0917<- data0917 %>% select(-c(STDORGCD09,STDORGCD17))

data0917$DSTRB09 <- cond09$DSTRBCD1[match(data0917$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0917$DSTRB17 <- cond17$DSTRBCD1[match(data0917$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data0917<- data0917 %>% filter(DSTRB09==0)
data0917<- data0917 %>% filter(DSTRB17==0)
data0917<- data0917 %>% select(-c(DSTRB09, DSTRB17))

data0917$TRT09 <- cond09$TRTCD1[match(data0917$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0917$TRT17 <- cond17$TRTCD1[match(data0917$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data0917<- data0917 %>% filter(TRT09==0)
data0917<- data0917 %>% filter(TRT17==0)
data0917<- data0917 %>% select(-c(TRT09, TRT17))

#2009-18
data0918$STDORGCD09 <- cond09$STDORGCD[match(data0918$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0918$STDORGCD18 <- cond18$STDORGCD[match(data0918$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data0918<- data0918 %>% filter(STDORGCD09==0)
data0918<- data0918 %>% filter(STDORGCD18==0)
data0918<- data0918 %>% select(-c(STDORGCD09,STDORGCD18))

data0918$DSTRB09 <- cond09$DSTRBCD1[match(data0918$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0918$DSTRB18 <- cond18$DSTRBCD1[match(data0918$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data0918<- data0918 %>% filter(DSTRB09==0)
data0918<- data0918 %>% filter(DSTRB18==0)
data0918<- data0918 %>% select(-c(DSTRB09, DSTRB18))

data0918$TRT09 <- cond09$TRTCD1[match(data0918$COUNTY_PLOT, cond09$COUNTY_PLOT)]
data0918$TRT18 <- cond18$TRTCD1[match(data0918$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data0918<- data0918 %>% filter(TRT09==0)
data0918<- data0918 %>% filter(TRT18==0)
data0918<- data0918 %>% select(-c(TRT09, TRT18))

#2010-15
data1015$STDORGCD10 <- cond10$STDORGCD[match(data1015$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1015$STDORGCD15 <- cond15$STDORGCD[match(data1015$COUNTY_PLOT, cond15$COUNTY_PLOT)]
data1015<- data1015 %>% filter(STDORGCD10==0)
data1015<- data1015 %>% filter(STDORGCD15==0)
data1015<- data1015 %>% select(-c(STDORGCD10,STDORGCD15))

data1015$DSTRB10 <- cond10$DSTRBCD1[match(data1015$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1015$DSTRB15 <- cond15$DSTRBCD1[match(data1015$COUNTY_PLOT, cond15$COUNTY_PLOT)]
data1015<- data1015 %>% filter(DSTRB10==0)
data1015<- data1015 %>% filter(DSTRB15==0)
data1015<- data1015 %>% select(-c(DSTRB10, DSTRB15))

data1015$TRT10 <- cond10$TRTCD1[match(data1015$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1015$TRT15 <- cond15$TRTCD1[match(data1015$COUNTY_PLOT, cond15$COUNTY_PLOT)]
data1015<- data1015 %>% filter(TRT10==0)
data1015<- data1015 %>% filter(TRT15==0)
data1015<- data1015 %>% select(-c(TRT10, TRT15))

#2010-16
data1016$STDORGCD10 <- cond10$STDORGCD[match(data1016$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1016$STDORGCD16 <- cond16$STDORGCD[match(data1016$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data1016<- data1016 %>% filter(STDORGCD10==0)
data1016<- data1016 %>% filter(STDORGCD16==0)
data1016<- data1016 %>% select(-c(STDORGCD10,STDORGCD16))

data1016$DSTRB10 <- cond10$DSTRBCD1[match(data1016$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1016$DSTRB16 <- cond16$DSTRBCD1[match(data1016$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data1016<- data1016 %>% filter(DSTRB10==0)
data1016<- data1016 %>% filter(DSTRB16==0)
data1016<- data1016 %>% select(-c(DSTRB10, DSTRB16))

data1016$TRT10 <- cond10$TRTCD1[match(data1016$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1016$TRT16 <- cond16$TRTCD1[match(data1016$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data1016<- data1016 %>% filter(TRT10==0)
data1016<- data1016 %>% filter(TRT16==0)
data1016<- data1016 %>% select(-c(TRT10, TRT16))

#2010-17
data1017$STDORGCD10 <- cond10$STDORGCD[match(data1017$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1017$STDORGCD17 <- cond17$STDORGCD[match(data1017$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data1017<- data1017 %>% filter(STDORGCD10==0)
data1017<- data1017 %>% filter(STDORGCD17==0)
data1017<- data1017 %>% select(-c(STDORGCD10,STDORGCD17))

data1017$DSTRB10 <- cond10$DSTRBCD1[match(data1017$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1017$DSTRB17 <- cond17$DSTRBCD1[match(data1017$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data1017<- data1017 %>% filter(DSTRB10==0)
data1017<- data1017 %>% filter(DSTRB17==0)
data1017<- data1017 %>% select(-c(DSTRB10, DSTRB17))

data1017$TRT10 <- cond10$TRTCD1[match(data1017$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1017$TRT17 <- cond17$TRTCD1[match(data1017$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data1017<- data1017 %>% filter(TRT10==0)
data1017<- data1017 %>% filter(TRT17==0)
data1017<- data1017 %>% select(-c(TRT10, TRT17))

#2010-18
data1018$STDORGCD10 <- cond10$STDORGCD[match(data1018$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1018$STDORGCD18 <- cond18$STDORGCD[match(data1018$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data1018<- data1018 %>% filter(STDORGCD10==0)
data1018<- data1018 %>% filter(STDORGCD18==0)
data1018<- data1018 %>% select(-c(STDORGCD10,STDORGCD18))

data1018$DSTRB10 <- cond10$DSTRBCD1[match(data1018$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1018$DSTRB18 <- cond18$DSTRBCD1[match(data1018$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data1018<- data1018 %>% filter(DSTRB10==0)
data1018<- data1018 %>% filter(DSTRB18==0)
data1018<- data1018 %>% select(-c(DSTRB10, DSTRB18))

data1018$TRT10 <- cond10$TRTCD1[match(data1018$COUNTY_PLOT, cond10$COUNTY_PLOT)]
data1018$TRT18 <- cond18$TRTCD1[match(data1018$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data1018<- data1018 %>% filter(TRT10==0)
data1018<- data1018 %>% filter(TRT18==0)
data1018<- data1018 %>% select(-c(TRT10, TRT18))

#2011-15
data1115$STDORGCD11 <- cond11$STDORGCD[match(data1115$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1115$STDORGCD15 <- cond15$STDORGCD[match(data1115$COUNTY_PLOT, cond15$COUNTY_PLOT)]
data1115<- data1115 %>% filter(STDORGCD11==0)
data1115<- data1115 %>% filter(STDORGCD15==0)
data1115<- data1115 %>% select(-c(STDORGCD11,STDORGCD15))

data1115$DSTRB11 <- cond11$DSTRBCD1[match(data1115$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1115$DSTRB15 <- cond15$DSTRBCD1[match(data1115$COUNTY_PLOT, cond15$COUNTY_PLOT)]
data1115<- data1115 %>% filter(DSTRB11==0)
data1115<- data1115 %>% filter(DSTRB15==0)
data1115<- data1115 %>% select(-c(DSTRB11, DSTRB15))

data1115$TRT11 <- cond11$TRTCD1[match(data1115$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1115$TRT15 <- cond15$TRTCD1[match(data1115$COUNTY_PLOT, cond15$COUNTY_PLOT)]
data1115<- data1115 %>% filter(TRT11==0)
data1115<- data1115 %>% filter(TRT15==0)
data1115<- data1115 %>% select(-c(TRT11, TRT15))

#2011-16
data1116$STDORGCD11 <- cond11$STDORGCD[match(data1116$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1116$STDORGCD16 <- cond16$STDORGCD[match(data1116$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data1116<- data1116 %>% filter(STDORGCD11==0)
data1116<- data1116 %>% filter(STDORGCD16==0)
data1116<- data1116 %>% select(-c(STDORGCD11,STDORGCD16))

data1116$DSTRB11 <- cond11$DSTRBCD1[match(data1116$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1116$DSTRB16 <- cond16$DSTRBCD1[match(data1116$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data1116<- data1116 %>% filter(DSTRB11==0)
data1116<- data1116 %>% filter(DSTRB16==0)
data1116<- data1116 %>% select(-c(DSTRB11, DSTRB16))

data1116$TRT11 <- cond11$TRTCD1[match(data1116$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1116$TRT16 <- cond16$TRTCD1[match(data1116$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data1116<- data1116 %>% filter(TRT11==0)
data1116<- data1116 %>% filter(TRT16==0)
data1116<- data1116 %>% select(-c(TRT11, TRT16))

#2011-17
data1117$STDORGCD11 <- cond11$STDORGCD[match(data1117$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1117$STDORGCD17 <- cond17$STDORGCD[match(data1117$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data1117<- data1117 %>% filter(STDORGCD11==0)
data1117<- data1117 %>% filter(STDORGCD17==0)
data1117<- data1117 %>% select(-c(STDORGCD11,STDORGCD17))

data1117$DSTRB11 <- cond11$DSTRBCD1[match(data1117$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1117$DSTRB17 <- cond17$DSTRBCD1[match(data1117$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data1117<- data1117 %>% filter(DSTRB11==0)
data1117<- data1117 %>% filter(DSTRB17==0)
data1117<- data1117 %>% select(-c(DSTRB11, DSTRB17))

data1117$TRT11 <- cond11$TRTCD1[match(data1117$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1117$TRT17 <- cond17$TRTCD1[match(data1117$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data1117<- data1117 %>% filter(TRT11==0)
data1117<- data1117 %>% filter(TRT17==0)
data1117<- data1117 %>% select(-c(TRT11, TRT17))

#2011-18
data1118$STDORGCD11 <- cond11$STDORGCD[match(data1118$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1118$STDORGCD18 <- cond18$STDORGCD[match(data1118$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data1118<- data1118 %>% filter(STDORGCD11==0)
data1118<- data1118 %>% filter(STDORGCD18==0)
data1118<- data1118 %>% select(-c(STDORGCD11,STDORGCD18))

data1118$DSTRB11 <- cond11$DSTRBCD1[match(data1118$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1118$DSTRB18 <- cond18$DSTRBCD1[match(data1118$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data1118<- data1118 %>% filter(DSTRB11==0)
data1118<- data1118 %>% filter(DSTRB18==0)
data1118<- data1118 %>% select(-c(DSTRB11, DSTRB18))

data1118$TRT11 <- cond11$TRTCD1[match(data1118$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1118$TRT18 <- cond18$TRTCD1[match(data1118$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data1118<- data1118 %>% filter(TRT11==0)
data1118<- data1118 %>% filter(TRT18==0)
data1118<- data1118 %>% select(-c(TRT11, TRT18))

#2011-19
data1119$STDORGCD11 <- cond11$STDORGCD[match(data1119$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1119$STDORGCD19 <- cond19$STDORGCD[match(data1119$COUNTY_PLOT, cond19$COUNTY_PLOT)]
data1119<- data1119 %>% filter(STDORGCD11==0)
data1119<- data1119 %>% filter(STDORGCD19==0)
data1119<- data1119 %>% select(-c(STDORGCD11,STDORGCD19))

data1119$DSTRB11 <- cond11$DSTRBCD1[match(data1119$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1119$DSTRB19 <- cond19$DSTRBCD1[match(data1119$COUNTY_PLOT, cond19$COUNTY_PLOT)]
data1119<- data1119 %>% filter(DSTRB11==0)
data1119<- data1119 %>% filter(DSTRB19==0)
data1119<- data1119 %>% select(-c(DSTRB11, DSTRB19))

data1119$TRT11 <- cond11$TRTCD1[match(data1119$COUNTY_PLOT, cond11$COUNTY_PLOT)]
data1119$TRT19 <- cond19$TRTCD1[match(data1119$COUNTY_PLOT, cond19$COUNTY_PLOT)]
data1119<- data1119 %>% filter(TRT11==0)
data1119<- data1119 %>% filter(TRT19==0)
data1119<- data1119 %>% select(-c(TRT11, TRT19))

#2012-16
data1216$STDORGCD12 <- cond12$STDORGCD[match(data1216$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1216$STDORGCD16 <- cond16$STDORGCD[match(data1216$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data1216<- data1216 %>% filter(STDORGCD12==0)
data1216<- data1216 %>% filter(STDORGCD16==0)
data1216<- data1216 %>% select(-c(STDORGCD12,STDORGCD16))

data1216$DSTRB12 <- cond12$DSTRBCD1[match(data1216$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1216$DSTRB16 <- cond16$DSTRBCD1[match(data1216$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data1216<- data1216 %>% filter(DSTRB12==0)
data1216<- data1216 %>% filter(DSTRB16==0)
data1216<- data1216 %>% select(-c(DSTRB12, DSTRB16))

data1216$TRT12 <- cond12$TRTCD1[match(data1216$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1216$TRT16 <- cond16$TRTCD1[match(data1216$COUNTY_PLOT, cond16$COUNTY_PLOT)]
data1216<- data1216 %>% filter(TRT12==0)
data1216<- data1216 %>% filter(TRT16==0)
data1216<- data1216 %>% select(-c(TRT12, TRT16))

#2012-17
data1217$STDORGCD12 <- cond12$STDORGCD[match(data1217$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1217$STDORGCD17 <- cond17$STDORGCD[match(data1217$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data1217<- data1217 %>% filter(STDORGCD12==0)
data1217<- data1217 %>% filter(STDORGCD17==0)
data1217<- data1217 %>% select(-c(STDORGCD12,STDORGCD17))

data1217$DSTRB12 <- cond12$DSTRBCD1[match(data1217$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1217$DSTRB17 <- cond17$DSTRBCD1[match(data1217$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data1217<- data1217 %>% filter(DSTRB12==0)
data1217<- data1217 %>% filter(DSTRB17==0)
data1217<- data1217 %>% select(-c(DSTRB12, DSTRB17))

data1217$TRT12 <- cond12$TRTCD1[match(data1217$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1217$TRT17 <- cond17$TRTCD1[match(data1217$COUNTY_PLOT, cond17$COUNTY_PLOT)]
data1217<- data1217 %>% filter(TRT12==0)
data1217<- data1217 %>% filter(TRT17==0)
data1217<- data1217 %>% select(-c(TRT12, TRT17))

#2012-18
data1218$STDORGCD12 <- cond12$STDORGCD[match(data1218$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1218$STDORGCD18 <- cond18$STDORGCD[match(data1218$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data1218<- data1218 %>% filter(STDORGCD12==0)
data1218<- data1218 %>% filter(STDORGCD18==0)
data1218<- data1218 %>% select(-c(STDORGCD12,STDORGCD18))

data1218$DSTRB12 <- cond12$DSTRBCD1[match(data1218$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1218$DSTRB18 <- cond18$DSTRBCD1[match(data1218$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data1218<- data1218 %>% filter(DSTRB12==0)
data1218<- data1218 %>% filter(DSTRB18==0)
data1218<- data1218 %>% select(-c(DSTRB12, DSTRB18))

data1218$TRT12 <- cond12$TRTCD1[match(data1218$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1218$TRT18 <- cond18$TRTCD1[match(data1218$COUNTY_PLOT, cond18$COUNTY_PLOT)]
data1218<- data1218 %>% filter(TRT12==0)
data1218<- data1218 %>% filter(TRT18==0)
data1218<- data1218 %>% select(-c(TRT12, TRT18))

#2012-19
data1219$STDORGCD12 <- cond12$STDORGCD[match(data1219$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1219$STDORGCD19 <- cond19$STDORGCD[match(data1219$COUNTY_PLOT, cond19$COUNTY_PLOT)]
data1219<- data1219 %>% filter(STDORGCD12==0)
data1219<- data1219 %>% filter(STDORGCD19==0)
data1219<- data1219 %>% select(-c(STDORGCD12,STDORGCD19))

data1219$DSTRB12 <- cond12$DSTRBCD1[match(data1219$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1219$DSTRB19 <- cond19$DSTRBCD1[match(data1219$COUNTY_PLOT, cond19$COUNTY_PLOT)]
data1219<- data1219 %>% filter(DSTRB12==0)
data1219<- data1219 %>% filter(DSTRB19==0)
data1219<- data1219 %>% select(-c(DSTRB12, DSTRB19))

data1219$TRT12 <- cond12$TRTCD1[match(data1219$COUNTY_PLOT, cond12$COUNTY_PLOT)]
data1219$TRT19 <- cond19$TRTCD1[match(data1219$COUNTY_PLOT, cond19$COUNTY_PLOT)]
data1219<- data1219 %>% filter(TRT12==0)
data1219<- data1219 %>% filter(TRT19==0)
data1219<- data1219 %>% select(-c(TRT12, TRT19))
```

## Biomass equation
#### log(biomass)=B0+B1*log(DIA)----> biomass= exp(BO+ B1*log(DIA))
β_0 and β_1 are determined using species’ taxonomic identity and wood specific gravity (green volume to dry-weight basis) (Chojnacky, Heath, and Jenkins 2014; Miles and Smith 2009). 
```{r}
biomass_function <- function(spp, dbh) { 
  case_when(
    spp== 43  ~ exp( -2.7765   +    2.4195 *log(dbh)), 
    spp== 67  ~ exp( -2.6327   +    2.4757 *log(dbh)),
    spp== 68  ~ exp( -2.6327   +    2.4757 *log(dbh)),
    spp== 110 ~ exp( -3.0506   +    2.6465 *log(dbh)),
    spp== 111 ~ exp( -3.0506   +    2.6465 *log(dbh)),
    spp== 115 ~ exp( -2.6177   +    2.4638 *log(dbh)),
    spp== 121 ~ exp( -3.0506   +    2.6465 *log(dbh)),
    spp== 131 ~ exp( -3.0506   +    2.6465 *log(dbh)),
    spp== 132 ~ exp( -3.0506   +    2.6465 *log(dbh)),
    spp== 221 ~ exp( -2.6327   +    2.4757 *log(dbh)),
    spp== 261 ~ exp( -2.3480   +    2.3876 *log(dbh)),
    spp== 311 ~ exp( -2.0470   +    2.3852 *log(dbh)),
    spp== 313 ~ exp( -2.0470   +    2.3852 *log(dbh)),
    spp== 316 ~ exp( -2.0470   +    2.3852 *log(dbh)),
    spp== 317 ~ exp( -2.0470   +    2.3852 *log(dbh)),
    spp== 318 ~ exp( -1.8011   +    2.3852 *log(dbh)),
    spp== 332 ~ exp( -2.4108   +    2.4177 *log(dbh)),
    spp== 341 ~ exp( -1.8011   +    2.3852 *log(dbh)),
    spp== 345 ~ exp( -2.5095   +    2.5437 *log(dbh)),
    spp== 356 ~ exp( -2.2118   +    2.4133 *log(dbh)), 
    spp== 367 ~ exp( -2.5497   +    2.5011 *log(dbh)),
    spp== 372 ~ exp( -2.2652   +    2.5349 *log(dbh)),
    spp== 373 ~ exp( -1.8096   +    2.3480 *log(dbh)), 
    spp== 391 ~ exp( -2.2652   +    2.5349 *log(dbh)),
    spp== 401 ~ exp( -2.5095   +    2.6175 *log(dbh)),
    spp== 402 ~ exp( -2.5095   +    2.6175 *log(dbh)),
    spp== 403 ~ exp( -2.5095   +    2.6175 *log(dbh)),
    spp== 404 ~ exp( -2.5095   +    2.6175 *log(dbh)),
    spp== 405 ~ exp( -2.5095   +    2.6175 *log(dbh)),
    spp== 407 ~ exp( -2.5095   +    2.6175 *log(dbh)),
    spp== 408 ~ exp( -2.5095   +    2.6175 *log(dbh)),
    spp== 409 ~ exp( -2.5095   +    2.6175 *log(dbh)),
    spp== 412 ~ exp( -2.5095   +    2.6175 *log(dbh)),
    spp== 413 ~ exp( -2.5095   +    2.6175 *log(dbh)),
    spp== 422 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 451 ~ exp( -2.0314   +    2.3524 *log(dbh)),
    spp== 461 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 462 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 471 ~ exp( -2.5095   +    2.5437 *log(dbh)),
    spp== 491 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 500 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 521 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 531 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 541 ~ exp( -1.8384   +    2.3524 *log(dbh)),
    spp== 544 ~ exp( -2.0314   +    2.3524 *log(dbh)),
    spp== 545 ~ exp( -2.0314   +    2.3524 *log(dbh)),
    spp== 552 ~ exp( -2.5095   +    2.5437 *log(dbh)),
    spp== 581 ~ exp( -2.0314   +    2.3524 *log(dbh)),
    spp== 591 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 601 ~ exp( -2.5095   +    2.5437 *log(dbh)),
    spp== 602 ~ exp( -2.5095   +    2.5437 *log(dbh)),
    spp== 611 ~ exp( -2.6390   +    2.5466 *log(dbh)),
    spp== 621 ~ exp( -2.5497   +    2.5011 *log(dbh)),
    spp== 641 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 651 ~ exp( -2.5497   +    2.5011 *log(dbh)),
    spp== 652 ~ exp( -2.5497   +    2.5011 *log(dbh)),
    spp== 653 ~ exp( -2.5497   +    2.5011 *log(dbh)),
    spp== 654 ~ exp( -2.5497   +    2.5011 *log(dbh)),
    spp== 658 ~ exp( -2.5497   +    2.5011 *log(dbh)),
    spp== 662 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 681 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 682 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 691 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 693 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 694 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 701 ~ exp( -2.2652   +    2.5349 *log(dbh)),
    spp== 711 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 712 ~ exp( -2.0314   +    2.3524 *log(dbh)),
    spp== 721 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 722 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 731 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 742 ~ exp( -2.4441   +    2.4561 *log(dbh)),
    spp== 760 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 762 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 766 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 802 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 806 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 808 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 812 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 813 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 819 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 820 ~ exp( -2.2198   +    2.4410 *log(dbh)),
    spp== 822 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 824 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 825 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 826 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 827 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 828 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 831 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 832 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 833 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 834 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 835 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 836 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 837 ~ exp( -2.0705   +    2.4410 *log(dbh)),
    spp== 838 ~ exp( -2.2198   +    2.4410 *log(dbh)),
    spp== 840 ~ exp( -2.2198   +    2.4410 *log(dbh)),
    spp== 841 ~ exp( -2.2198   +    2.4410 *log(dbh)),
    spp== 842 ~ exp( -2.2198   +    2.4410 *log(dbh)),
    spp== 858 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 901 ~ exp( -2.5095   +    2.5437 *log(dbh)),
    spp== 922 ~ exp( -2.4441   +    2.4561 *log(dbh)),
    spp== 931 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 951 ~ exp( -2.4108   +    2.4177 *log(dbh)),
    spp== 952 ~ exp( -2.4108   +    2.4177 *log(dbh)),
    spp== 953 ~ exp( -2.4108   +    2.4177 *log(dbh)),
    spp== 971 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 972 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 975 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 976 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 993 ~ exp( -1.8011   +    2.3852 *log(dbh)),
    spp== 994 ~ exp( -2.2118   +    2.4133 *log(dbh)),
    spp== 996 ~ exp( -1.8011   +    2.3852 *log(dbh)),)}
```

### now applying that equation
```{r}
#2009-14
bio0914<- data0914 %>% mutate(bio09 = biomass_function(spp = SPCD, dbh = DIA09))
bio0914<- bio0914 %>% mutate(bio14 = biomass_function(spp = SPCD, dbh = DIA14))

#2009-15
bio0915<- data0915 %>% mutate(bio09 = biomass_function(spp = SPCD, dbh = DIA09))
bio0915<- bio0915 %>% mutate(bio15 = biomass_function(spp = SPCD, dbh = DIA15))

#2009-16
bio0916<- data0916 %>% mutate(bio09 = biomass_function(spp = SPCD, dbh = DIA09))
bio0916<- bio0916 %>% mutate(bio16 = biomass_function(spp = SPCD, dbh = DIA16))

#2009-17
bio0917<- data0917 %>% mutate(bio09 = biomass_function(spp = SPCD, dbh = DIA09))
bio0917<- bio0917 %>% mutate(bio17 = biomass_function(spp = SPCD, dbh = DIA17))

#2009-18
bio0918<- data0918 %>% mutate(bio09 = biomass_function(spp = SPCD, dbh = DIA09))
bio0918<- bio0918 %>% mutate(bio18 = biomass_function(spp = SPCD, dbh = DIA18))

#2010-15
bio1015<- data1015 %>% mutate(bio10 = biomass_function(spp = SPCD, dbh = DIA10))
bio1015<- bio1015 %>% mutate(bio15 = biomass_function(spp = SPCD, dbh = DIA15))

#2010-16
bio1016<- data1016 %>% mutate(bio10 = biomass_function(spp = SPCD, dbh = DIA10))
bio1016<- bio1016 %>% mutate(bio16 = biomass_function(spp = SPCD, dbh = DIA16))

#2010-17
bio1017<- data1017 %>% mutate(bio10 = biomass_function(spp = SPCD, dbh = DIA10))
bio1017<- bio1017 %>% mutate(bio17 = biomass_function(spp = SPCD, dbh = DIA17))

#2010-18
bio1018<- data1018 %>% mutate(bio10 = biomass_function(spp = SPCD, dbh = DIA10))
bio1018<- bio1018 %>% mutate(bio18 = biomass_function(spp = SPCD, dbh = DIA18))

#2011-15
bio1115<- data1115 %>% mutate(bio11 = biomass_function(spp = SPCD, dbh = DIA11))
bio1115<- bio1115 %>% mutate(bio15 = biomass_function(spp = SPCD, dbh = DIA15))

#2011-16
bio1116<- data1116 %>% mutate(bio11 = biomass_function(spp = SPCD, dbh = DIA11))
bio1116<- bio1116 %>% mutate(bio16 = biomass_function(spp = SPCD, dbh = DIA16))

#2011-17
bio1117<- data1117 %>% mutate(bio11 = biomass_function(spp = SPCD, dbh = DIA11))
bio1117<- bio1117 %>% mutate(bio17 = biomass_function(spp = SPCD, dbh = DIA17))

#2011-18
bio1118<- data1118 %>% mutate(bio11 = biomass_function(spp = SPCD, dbh = DIA11))
bio1118<- bio1118 %>% mutate(bio18 = biomass_function(spp = SPCD, dbh = DIA18))

#2011-19
bio1119<- data1119 %>% mutate(bio11 = biomass_function(spp = SPCD, dbh = DIA11))
bio1119<- bio1119 %>% mutate(bio19 = biomass_function(spp = SPCD, dbh = DIA19))

#2012-16
bio1216<- data1216 %>% mutate(bio12 = biomass_function(spp = SPCD, dbh = DIA12))
bio1216<- bio1216 %>% mutate(bio16 = biomass_function(spp = SPCD, dbh = DIA16))

#2012-17
bio1217<- data1217 %>% mutate(bio12 = biomass_function(spp = SPCD, dbh = DIA12))
bio1217<- bio1217 %>% mutate(bio17 = biomass_function(spp = SPCD, dbh = DIA17))

#2012-18
bio1218<- data1218 %>% mutate(bio12 = biomass_function(spp = SPCD, dbh = DIA12))
bio1218<- bio1218 %>% mutate(bio18 = biomass_function(spp = SPCD, dbh = DIA18))

#2012-19
bio1219<- data1219 %>% mutate(bio12 = biomass_function(spp = SPCD, dbh = DIA12))
bio1219<- bio1219 %>% mutate(bio19 = biomass_function(spp = SPCD, dbh = DIA19))
```

### adding some necessary vars (subplot label and tpha)
```{r}
#adding subplot label
bio0914$COUNTY_PLOT_SUB <- paste(bio0914$COUNTY_PLOT,bio0914$SUBP, sep="_")
bio0915$COUNTY_PLOT_SUB <- paste(bio0915$COUNTY_PLOT,bio0915$SUBP, sep="_")
bio0916$COUNTY_PLOT_SUB <- paste(bio0916$COUNTY_PLOT,bio0916$SUBP, sep="_")
bio0917$COUNTY_PLOT_SUB <- paste(bio0917$COUNTY_PLOT,bio0917$SUBP, sep="_")
bio0918$COUNTY_PLOT_SUB <- paste(bio0918$COUNTY_PLOT,bio0918$SUBP, sep="_")
bio1015$COUNTY_PLOT_SUB <- paste(bio1015$COUNTY_PLOT,bio1015$SUBP, sep="_")
bio1016$COUNTY_PLOT_SUB <- paste(bio1016$COUNTY_PLOT,bio1016$SUBP, sep="_")
bio1017$COUNTY_PLOT_SUB <- paste(bio1017$COUNTY_PLOT,bio1017$SUBP, sep="_")
bio1018$COUNTY_PLOT_SUB <- paste(bio1018$COUNTY_PLOT,bio1018$SUBP, sep="_")
bio1115$COUNTY_PLOT_SUB <- paste(bio1115$COUNTY_PLOT,bio1115$SUBP, sep="_")
bio1116$COUNTY_PLOT_SUB <- paste(bio1116$COUNTY_PLOT,bio1116$SUBP, sep="_")
bio1117$COUNTY_PLOT_SUB <- paste(bio1117$COUNTY_PLOT,bio1117$SUBP, sep="_")
bio1118$COUNTY_PLOT_SUB <- paste(bio1118$COUNTY_PLOT,bio1118$SUBP, sep="_")
bio1119$COUNTY_PLOT_SUB <- paste(bio1119$COUNTY_PLOT,bio1119$SUBP, sep="_")
bio1216$COUNTY_PLOT_SUB <- paste(bio1216$COUNTY_PLOT,bio1216$SUBP, sep="_")
bio1217$COUNTY_PLOT_SUB <- paste(bio1217$COUNTY_PLOT,bio1217$SUBP, sep="_")
bio1218$COUNTY_PLOT_SUB <- paste(bio1218$COUNTY_PLOT,bio1218$SUBP, sep="_")
bio1219$COUNTY_PLOT_SUB <- paste(bio1219$COUNTY_PLOT,bio1219$SUBP, sep="_")

#trees/ha
bio0914<-bio0914 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio0915<-bio0915 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio0916<-bio0916 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio0917<-bio0917 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio0918<-bio0918 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1015<-bio1015 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1016<-bio1016 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1017<-bio1017 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1018<-bio1018 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1115<-bio1115 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1116<-bio1116 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1117<-bio1117 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1118<-bio1118 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1119<-bio1119 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1216<-bio1216 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1217<-bio1217 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1218<-bio1218 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
bio1219<-bio1219 %>%  group_by(COUNTY_PLOT_SUB) %>% mutate(tpha=n()/0.0168)
```

### now getting FIA subplot (called plot in the manuscript) sums of biomass
```{r}
#2009-14
bio0914<- bio0914 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio09=sum(bio09,na.rm=T),sumbio14=sum(bio14,na.rm = T))
bio0914$bio_change <- ((bio0914$sumbio14- bio0914$sumbio09)/5)
#2009-15
bio0915<- bio0915 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio09=sum(bio09,na.rm=T),sumbio15=sum(bio15,na.rm=T))
bio0915$bio_change <- ((bio0915$sumbio15- bio0915$sumbio09)/6)
#2009-16
bio0916<- bio0916 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio09=sum(bio09,na.rm=T),sumbio16=sum(bio16,na.rm=T))
bio0916$bio_change <- ((bio0916$sumbio16- bio0916$sumbio09)/7)
#2009-17
bio0917<- bio0917 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio09=sum(bio09,na.rm=T),sumbio17=sum(bio17,na.rm=T))
bio0917$bio_change <- ((bio0917$sumbio17- bio0917$sumbio09)/8)
#2009-18
bio0918<- bio0918 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio09=sum(bio09,na.rm=T),sumbio18=sum(bio18,na.rm=T))
bio0918$bio_change <- ((bio0918$sumbio18- bio0918$sumbio09)/9)
#2010-15
bio1015<- bio1015 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio10=sum(bio10,na.rm=T),sumbio15=sum(bio15,na.rm=T))
bio1015$bio_change <- ((bio1015$sumbio15- bio1015$sumbio10)/5)
#2010-16
bio1016<- bio1016 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio10=sum(bio10,na.rm=T),sumbio16=sum(bio16,na.rm=T))
bio1016$bio_change <- ((bio1016$sumbio16- bio1016$sumbio10)/6)
#2010-17
bio1017<- bio1017 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio10=sum(bio10,na.rm=T),sumbio17=sum(bio17,na.rm=T))
bio1017$bio_change <- ((bio1017$sumbio17- bio1017$sumbio10)/7)
#2010-18
bio1018<- bio1018 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio10=sum(bio10,na.rm=T),sumbio18=sum(bio18,na.rm=T))
bio1018$bio_change <- ((bio1018$sumbio18- bio1018$sumbio10)/8)
#2011-15
bio1115<- bio1115 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio11=sum(bio11,na.rm=T),sumbio15=sum(bio15,na.rm=T))
bio1115$bio_change <- ((bio1115$sumbio15- bio1115$sumbio11)/4)
#2011-16
bio1116<- bio1116 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio11=sum(bio11,na.rm=T),sumbio16=sum(bio16,na.rm=T))
bio1116$bio_change <- ((bio1116$sumbio16- bio1116$sumbio11)/5)
#2011-17
bio1117<- bio1117 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio11=sum(bio11,na.rm=T),sumbio17=sum(bio17,na.rm=T))
bio1117$bio_change <- ((bio1117$sumbio17- bio1117$sumbio11)/6)
#2011-18
bio1118<- bio1118 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio11=sum(bio11,na.rm=T),sumbio18=sum(bio18,na.rm=T))
bio1118$bio_change <- ((bio1118$sumbio18- bio1118$sumbio11)/7)
#2011-19
bio1119<- bio1119 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio11=sum(bio11,na.rm=T),sumbio19=sum(bio19,na.rm=T))
bio1119$bio_change <- ((bio1119$sumbio19- bio1119$sumbio11)/8)
#2012-16
bio1216<- bio1216 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio12=sum(bio12,na.rm=T),sumbio16=sum(bio16,na.rm=T))
bio1216$bio_change <- ((bio1216$sumbio16- bio1216$sumbio12)/4)
#2012-17
bio1217<- bio1217 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio12=sum(bio12,na.rm=T),sumbio17=sum(bio17,na.rm=T))
bio1217$bio_change <- ((bio1217$sumbio17- bio1217$sumbio12)/5)
#2012-18
bio1218<- bio1218 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio12=sum(bio12,na.rm=T),sumbio18=sum(bio18,na.rm=T))
bio1218$bio_change <- ((bio1218$sumbio18- bio1218$sumbio12)/6)
#2012-19
bio1219<- bio1219 %>% group_by(COUNTY_PLOT_SUB) %>% mutate(sumbio12=sum(bio12,na.rm=T),sumbio19=sum(bio19,na.rm=T))
bio1219$bio_change <- ((bio1219$sumbio19- bio1219$sumbio12)/7)
```

### now to merge all years and plots
```{r}
mer0914 <- bio0914 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer0915 <- bio0915 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer0916 <- bio0916 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer0917 <- bio0917 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer0918 <- bio0918 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1015 <- bio1015 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1016 <- bio1016 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1017 <- bio1017 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1018 <- bio1018 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1115 <- bio1115 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1116 <- bio1116 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1117 <- bio1117 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1118 <- bio1118 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1119 <- bio1119 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1216 <- bio1216 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1217 <- bio1217 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1218 <- bio1218 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")
mer1219 <- bio1219 %>% select('COUNTY_PLOT_SUB', 'bio_change','SPCD','STDAGE','BA',"COUNTY_PLOT","tpha")


#now for the complete merge
compdata_AL<- mer0916 %>% bind_rows(mer0914,mer0915,mer0917, mer0918, mer1015,mer1016,mer1017, mer1018, mer1115,mer1116, mer1117, mer1118, mer1119, mer1216, mer1217, mer1218, mer1219)
```

#### converting to plot level biomass and removing unidentified species 
```{r}
compdata_AL$bio_change<- compdata_AL$bio_change/0.0168

compdata_AL<- compdata_AL %>% filter(SPCD != "998")
compdata_AL<- compdata_AL %>% filter(SPCD != "999")
```

## calculating species richness (S) and shannon-weiner's diversity index (H)
species richness
```{r}
compdata_AL<- compdata_AL %>% group_by(COUNTY_PLOT_SUB) %>% mutate(S = n_distinct(SPCD)) %>% ungroup()
```
Shannon-Weiner's diversity index
```{r message=FALSE, warning=FALSE}
compdata_AL$CNTY_PLT_SUB_SPCD<-paste(compdata_AL$COUNTY_PLOT_SUB,compdata_AL$SPCD,sep="#")
#lets compute shannon's index
shdata<- compdata_AL %>% select(c(COUNTY_PLOT_SUB,SPCD,S,bio_change))
#total individuals
shdata<- shdata %>% group_by(COUNTY_PLOT_SUB) %>% mutate(totind = n()) %>% ungroup()
#number of each species
shdata<- shdata %>% group_by(COUNTY_PLOT_SUB, SPCD) %>% mutate(indsp = n()) %>% ungroup()
#ratio
shdata<- shdata %>% mutate(p = indsp/totind)
#pln(p)
shdata<- shdata %>% mutate(plnp = p*log(p))
#sum1
shdata<- shdata %>% group_by(COUNTY_PLOT_SUB,SPCD) %>% summarise(sum1 = mean(plnp))%>% ungroup()
#final sum
shdata<- shdata %>% group_by(COUNTY_PLOT_SUB) %>% mutate(summ= sum(sum1)) %>% ungroup()
#final step
shdata<- shdata %>% mutate(H= -1*summ)
compdata_AL$H <- shdata$H[match(compdata_AL$COUNTY_PLOT_SUB,shdata$COUNTY_PLOT_SUB)]
```

#### we also looked at using basal area to calculate H but decided against using it, here is the code for that calculation anyways
```{r}
# making proportions based on BA 
shdata<- compdata_AL %>% select(c(COUNTY_PLOT_SUB,SPCD,CNTY_PLT_SUB_SPCD,BA))
#species within plot BA
shdata<- shdata %>% group_by(CNTY_PLT_SUB_SPCD) %>% mutate(spsum= sum(BA))
#totalplot BA
shdata<- shdata %>% group_by(COUNTY_PLOT_SUB) %>% mutate(pltsum = sum(BA))
#proportion
shdata<- shdata %>% mutate(p= spsum/pltsum)
compdata_AL$prop <- shdata$p[match(compdata_AL$CNTY_PLT_SUB_SPCD,shdata$CNTY_PLT_SUB_SPCD)]

#now applying Shannons using that proportion
#pln(p)
shdata<- shdata %>% mutate(plnp = p*log(p))
shdata<- shdata %>% filter(!duplicated(CNTY_PLT_SUB_SPCD)) 
shdata<- shdata %>% group_by(COUNTY_PLOT_SUB)%>%  mutate(H=-1*sum(plnp))

compdata_AL$HBA <- shdata$H[match(compdata_AL$COUNTY_PLOT_SUB, shdata$COUNTY_PLOT_SUB)]

```

now to remove duplicated lines and negative biomass change
```{r}
compdata_AL<- compdata_AL %>% filter(!duplicated(CNTY_PLT_SUB_SPCD)) %>% filter(bio_change>0)
```

### including environmental variables
```{r}
#from COND table
compdata_AL$slope<- cond_tab$SLOPE[match(compdata_AL$COUNTY_PLOT, cond_tab$COUNTY_PLOT)]
compdata_AL$slope<- atan(compdata_AL$slope/100)

#from PLOT table
compdata_AL$LAT <- plot_tab$LAT[match(compdata_AL$COUNTY_PLOT, plot_tab$COUNTY_PLOT)]
compdata_AL$ELEV<-plot_tab$ELEV[match(compdata_AL$COUNTY_PLOT, plot_tab$COUNTY_PLOT)]
```

labeling data as being from Alabama (important for future merge to region)
```{r}
compdata_AL$STATECD=1
```

#### adding functional traits and exporting
```{r}
#FROM FUNC DATA
FUNC_GRP <- read.csv("namedata.csv",header = T)
compdata_AL$phenology<- FUNC_GRP$phenology[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$n_fixer<- FUNC_GRP$N_fixer[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$specific_gravity<- FUNC_GRP$specific_gravity[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$maxheight<- FUNC_GRP$max_height_m[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$fruit_type<- FUNC_GRP$fruit_type[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$fruit_length<- FUNC_GRP$fruit_length_cm[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$growth_rate<- FUNC_GRP$growth_rate[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$life_span<- FUNC_GRP$life_span[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$shade_tol<- FUNC_GRP$shade_tolerance[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$drought_tol<- FUNC_GRP$drought_tolerance[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$fire_tol<- FUNC_GRP$fire_tolerance[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]
compdata_AL$anaerobic_tol<- FUNC_GRP$anaerobic_tolerance[match(compdata_AL$SPCD, FUNC_GRP$SPCD)]

# exporting functional traits and biomass ---------------------------------
write.csv(compdata_AL,"C:\\Users\\elbaa\\OneDrive\\Desktop\\compdata_AL.csv")
```

