---
title: "rFIA remote markdown"
author: "Elizabeth Baach"
date: "2/10/2021"
output: 
  html_document: 
    keep_md: yes
---
first we need to load the packages
```{r}
library(rFIA)
library(dplyr)
library(tidyverse)
```
#then we can remotely add MS FIA data
use this to download... but if already downloaded
```{r}
getFIA(c('MS'), dir = 'C:/Users/elbaa/OneDrive/Desktop/gitMSDATA', load = FALSE)

```
use this to load **preferred**
```{r}
MS <- readFIA('C:/Users/elbaa/OneDrive/Desktop/gitMSDATA', inMemory = FALSE)
```

#next, lets run the rFIA functions
first lets define our datbase as MS
```{r}
#must first define db at fia 
biomass(db = MS)
```
filtering using natural stand origin we get:
```{r}
originbio<- biomass(MS, byPlot = TRUE, grpBy = c("STDORGCD")) %>% filter(STDORGCD ==0)
origindiv<- diversity(MS, byPlot = TRUE, grpBy = "STDORGCD")%>% filter(STDORGCD ==0)
originvit<- vitalRates(MS, byPlot = TRUE, grpBy = "STDORGCD")%>% filter(STDORGCD ==0)
```
great! now we can add our favorite columns from the rFIA functions we just ran to a common table lets call it adddata
```{r}
adddata$S <- origindiv$S[match(adddata$pltID, origindiv$pltID)]
adddata$H<- origindiv$H[match(adddata$pltID, origindiv$pltID)]
adddata$BIO_GROW<- originvit$BIO_GROW[match(adddata$pltID,originvit$pltID)]
adddata$BIO_GROW_AC<- originvit$BIO_GROW_AC[match(adddata$pltID, originvit$pltID)]
```

#sweet! now we want to add columns from the general FIA database
first we need to load those tables remotely
```{r}
TREE <- readFIA(dir = 'C:/Users/elbaa/OneDrive/Desktop/gitMSDATA', tables = 'TREE', inMemory = TRUE)
COND <- readFIA(dir = 'C:/Users/elbaa/OneDrive/Desktop/gitMSDATA', tables = 'COND', inMemory = TRUE)

```
nice! now we can add our favorite columns from TREE and COND to adddata so everything we want is in one place
lets start with tree
```{r}
adddata$SPCD<- TREE$TREE$SPCD[match(adddata$PLT_CN, TREE$TREE$PLT_CN)]
adddata$COUNTYCD<- TREE$TREE$COUNTYCD[match(adddata$PLT_CN, TREE$TREE$PLT_CN)]
adddata$UNITCD <- TREE$TREE$UNITCD[match(adddata$PLT_CN, TREE$TREE$PLT_CN)]
adddata$DRYBIO_AG <- TREE$TREE$DRYBIO_AG[match(adddata$PLT_CN, TREE$TREE$PLT_CN)]
```
and heres from cond
```{r}
adddata$FORTYPCD <- COND$COND$FORTYPCD[match(adddata$PLT_CN, COND$COND$PLT_CN)]

```
awesome! now lets see if we can make the spread function work to see biomass overtime
```{r}
newtest<- adddata %>% group_by(pltID,YEAR) %>% summarise(changebiomass= sum(DRYBIO_AG))
newtest<- newtest %>% filter(biomass>0)
spdtest<- spread(newtest, YEAR, changebiomass, fill = NA)
view(spdtest)
#this part below isnt working 
# spdtest$mct <- rowSums(is.na(spdtest))
# spdtest<- spdtest %>% filter(mct<10)

```

