---
title: "rFIA remote markdown"
author: "Elizabeth Baach"
date: "2/10/2021"
output: 
  html_document: 
    keep_md: yes
---
first we need to load the packages
```{r}
library(rFIA)
library(dplyr)
library(tidyverse)
```
#then we can remotely add MS FIA data
use this to download... but if already downloaded
```{r}
getFIA(c('MS'), dir = 'name', load = FALSE)

```
use this to load preferred
```{r}
MS <- readFIA('C:/Users/elbaa/OneDrive/Desktop/gitMSDATA', inMemory = FALSE)
```

#next, lets run the rFIA functions
first lets define our datbase as MS
```{r}
#must first define db at fia 
biomass(db = MS)
```
filtering using natural stand origin we get:
```{r}
originbio<- biomass(MS, byPlot = TRUE, grpBy = c("STDORGCD")) %>% filter(STDORGCD ==0)
origindiv<- diversity(MS, byPlot = TRUE, grpBy = "STDORGCD")%>% filter(STDORGCD ==0)
originvit<- vitalRates(MS, byPlot = TRUE, grpBy = "STDORGCD")%>% filter(STDORGCD ==0)
```
great! now we can add our favorite columns from the rFIA functions we just ran to a common table lets call it adddata
```{r}
adddata<- originbio
adddata$S <- origindiv$S[match(adddata$pltID, origindiv$pltID)]
adddata$H <- origindiv$H[match(adddata$pltID, origindiv$pltID)]
adddata$BIO_GROW<- originvit$BIO_GROW[match(adddata$pltID,originvit$pltID)]
adddata$BIO_GROW_AC<- originvit$BIO_GROW_AC[match(adddata$pltID, originvit$pltID)]
```

#sweet! now we want to add columns from the general FIA database
first we need to load those tables remotely
```{r}
TREE <- readFIA(dir = 'C:/Users/elbaa/OneDrive/Desktop/gitMSDATA', tables = 'TREE', inMemory = TRUE)
COND <- readFIA(dir = 'C:/Users/elbaa/OneDrive/Desktop/gitMSDATA', tables = 'COND', inMemory = TRUE)

```
nice! now we can add our favorite columns from TREE and COND to adddata so everything we want is in one place
lets start with tree
```{r}
adddata$SPCD<- TREE$TREE$SPCD[match(adddata$PLT_CN, TREE$TREE$PLT_CN)]
adddata$COUNTYCD<- TREE$TREE$COUNTYCD[match(adddata$PLT_CN, TREE$TREE$PLT_CN)]
adddata$UNITCD <- TREE$TREE$UNITCD[match(adddata$PLT_CN, TREE$TREE$PLT_CN)]
adddata$DRYBIO_AG <- TREE$TREE$DRYBIO_AG[match(adddata$PLT_CN, TREE$TREE$PLT_CN)]
```
and heres from cond
```{r}
adddata$FORTYPCD <- COND$COND$FORTYPCD[match(adddata$PLT_CN, COND$COND$PLT_CN)]

```
awesome! now lets see if we can make the spread function work to see biomass overtime
```{r}
newtest<- adddata %>% group_by(pltID,YEAR) %>% summarise(biomass=sum(DRYBIO_AG))
newtest<- newtest %>% filter(biomass>0)
spdtest<- spread(newtest, YEAR, biomass, fill = NA)
view(spdtest)

```
ok, nows the time to get some subtraction going
first lets select only 2009-2019, then name them so we can use them properly
```{r}
data<- spdtest  %>% select(-c("2006","2007","2008"))
data<- rename(data, c(X2019="2019", X2018="2018", X2017="2017", X2016="2016", X2015="2015", X2014="2014", X2013="2013", X2012="2012", X2011="2011", X2010="2010", X2009="2009"))
```
sweet! Now time for the monster if_else section
```{r}
data$diff1918=   if_else((data$X2019>0)&(data$X2018>0),(data$X2019-data$X2018)/1,0) 
data$diff1917=  if_else((data$X2019>0)&(data$X2017>0),(data$X2019-data$X2017)/2,0)
data$diff1916=  if_else((data$X2019>0)&(data$X2016>0),(data$X2019-data$X2016)/3,0)
data$diff1915=  if_else((data$X2019>0)&(data$X2015>0),(data$X2019-data$X2015)/4,0)
data$diff1914= if_else((data$X2019>0)&(data$X2014>0),(data$X2019-data$X2014)/5,0)
data$diff1913=  if_else((data$X2019>0)&(data$X2013>0),(data$X2019-data$X2013)/6,0)
data$diff1912=  if_else((data$X2019>0)&(data$X2012>0),(data$X2019-data$X2012)/7,0)
data$diff1911=  if_else((data$X2019>0)&(data$X2011>0),(data$X2019-data$X2011)/8,0)
data$diff1910=  if_else((data$X2019>0)&(data$X2010>0),(data$X2019-data$X2010)/9,0)
data$diff1909=  if_else((data$X2019>0)&(data$X2009>0),(data$X2019-data$X2009)/10,0)

data$diff1817=             if_else((data$X2018>0)&(data$X2017>0),(data$X2018-data$X2017)/1,0)
data$diff1816=             if_else((data$X2018>0)&(data$X2016>0),(data$X2018-data$X2016)/2,0)
data$diff1815=             if_else((data$X2018>0)&(data$X2015>0),(data$X2018-data$X2015)/3,0)
data$diff1814=          if_else((data$X2018>0)&(data$X2014>0),(data$X2018-data$X2014)/4,0)
data$diff1813=             if_else((data$X2018>0)&(data$X2013>0),(data$X2018-data$X2013)/5,0)
data$diff1812=       if_else((data$X2018>0)&(data$X2012>0),(data$X2018-data$X2012)/6,0)
data$diff1811=         if_else((data$X2018>0)&(data$X2011>0),(data$X2018-data$X2011)/7,0)
data$diff1810=            if_else((data$X2018>0)&(data$X2010>0),(data$X2018-data$X2010)/8,0)
data$diff1809=            if_else((data$X2018>0)&(data$X2009>0),(data$X2018-data$X2009)/9,0)

data$diff1716=            if_else((data$X2017>0)&(data$X2016>0),(data$X2017-data$X2016)/1,0)
data$diff1715=            if_else((data$X2017>0)&(data$X2015>0),(data$X2017-data$X2015)/2,0)
data$diff1714=            if_else((data$X2017>0)&(data$X2014>0),(data$X2017-data$X2014)/3,0)
data$diff1713=           if_else((data$X2017>0)&(data$X2013>0),(data$X2017-data$X2013)/4,0)
data$diff1712=       if_else((data$X2017>0)&(data$X2012>0),(data$X2017-data$X2012)/5,0)
data$diff1711=        if_else((data$X2017>0)&(data$X2011>0),(data$X2017-data$X2011)/6,0)
data$diff1710=        if_else((data$X2017>0)&(data$X2010>0),(data$X2017-data$X2010)/7,0)
data$diff1709=          if_else((data$X2017>0)&(data$X2009>0),(data$X2017-data$X2009)/8,0)

data$diff1615=      if_else((data$X2016>0)&(data$X2015>0),(data$X2016-data$X2015)/1,0)
data$diff1614=   if_else((data$X2016>0)&(data$X2014>0),(data$X2016-data$X2014)/2,0)
data$diff1613=    if_else((data$X2016>0)&(data$X2013>0),(data$X2016-data$X2013)/3,0)
data$diff1612=    if_else((data$X2016>0)&(data$X2012>0),(data$X2016-data$X2012)/4,0)
data$diff1611=    if_else((data$X2016>0)&(data$X2011>0),(data$X2016-data$X2011)/5,0)
data$diff1610=       if_else((data$X2016>0)&(data$X2010>0),(data$X2016-data$X2010)/6,0)
data$diff1609=    if_else((data$X2016>0)&(data$X2009>0),(data$X2016-data$X2009)/7,0)

data$diff1514=        if_else((data$X2015>0)&(data$X2014>0),(data$X2015-data$X2014)/1,0)
data$diff1513=         if_else((data$X2015>0)&(data$X2013>0),(data$X2015-data$X2013)/2,0)
data$diff1512=        if_else((data$X2015>0)&(data$X2012>0),(data$X2015-data$X2012)/3,0)
data$diff1511=        if_else((data$X2015>0)&(data$X2011>0),(data$X2015-data$X2011)/4,0)
data$diff1510=       if_else((data$X2015>0)&(data$X2010>0),(data$X2015-data$X2010)/5,0)
data$diff1509=       if_else((data$X2015>0)&(data$X2009>0),(data$X2015-data$X2009)/6,0)

data$diff1413=       if_else((data$X2014>0)&(data$X2013>0),(data$X2014-data$X2013)/1,0)
data$diff1412=       if_else((data$X2014>0)&(data$X2012>0),(data$X2014-data$X2012)/2,0)
data$diff1411=       if_else((data$X2014>0)&(data$X2011>0),(data$X2014-data$X2011)/3,0)
data$diff1410=       if_else((data$X2014>0)&(data$X2010>0),(data$X2014-data$X2010)/4,0)
data$diff1409=       if_else((data$X2014>0)&(data$X2009>0),(data$X2014-data$X2009)/5,0)

data$diff1312=       if_else((data$X2013>0)&(data$X2012>0),(data$X2013-data$X2012)/1,0)
data$diff1311=        if_else((data$X2013>0)&(data$X2011>0),(data$X2013-data$X2011)/2,0)
data$diff1310=         if_else((data$X2013>0)&(data$X2010>0),(data$X2013-data$X2010)/3,0)
data$diff1309=         if_else((data$X2013>0)&(data$X2009>0),(data$X2013-data$X2009)/4,0)

data$diff1211=          if_else((data$X2012>0)&(data$X2011>0),(data$X2012-data$X2011)/1,0)
data$diff1210=          if_else((data$X2012>0)&(data$X2010>0),(data$X2012-data$X2010)/2,0)
data$diff1209=           if_else((data$X2012>0)&(data$X2009>0),(data$X2012-data$X2009)/3,0)

data$diff1110=       if_else((data$X2011>0)&(data$X2010>0),(data$X2011-data$X2010)/1,0)
data$diff1109=       if_else((data$X2011>0)&(data$X2009>0),(data$X2011-data$X2009)/2,0)

data$diff1009=        if_else((data$X2010>0)&(data$X2009>0),(data$X2010-data$X2009)/1,0)


```

very good! now that thats done, lets make a new table to work with and make some changes
```{r}
dataavg<- data
dataavg$mean <- rowMeans(dataavg[,16:69], na.rm=TRUE)
dataavg <- dataavg %>% select("mean", everything())
summary(dataavg$mean)
```

awesome! moment of truth... lets make a scatterplot!
```{r}
datasum<- dataavg %>% group_by(pltID) %>% summarise(changeinbiomass=mean)
ggplot(datasum, aes(x= pltID, y= changeinbiomass))+ geom_point()
```

